Online {\IterComp}

### 2017-11-15 results.tex
minor comments
instrumentation evaluation rewrite

%In this section we discuss our experimental evaluation. Afterwards, we discuss our results concerning the instrumentations for profiling
%our work metric. Finally we present the results of the online {\itercomp}.

@@

@@

In this experiment we evaluate the runtime overhead introduced by the work efficiency profiling. We compile the program using the
\texttt{-O3}. We then measure the wall-clock time of the instrumented and the native versions to calculate the runtime overhead. For
each benchmark, we compute the average overhead over all its 1,000 inputs.
Figure~\ref{fig:overhead-O3} shows the runtime overhead imposed by the work profiling using each of the proposed work profiling strategies over
their non-instrumented counterparts.
%The WPO relaxation was performed using profiling information for every specific input, which provides
%the perfect information to perform the WPO relaxation. This means that the experiments show its best performance enabled by having perfect
%profiling information.

%\begin{figure}[t]
%    \centering
%    \includegraphics[width=0.395\textwidth]{figs/adpcm_d-cfg-instr.pdf}
%    \caption{CFG of the function that contains the hot loop of the {\flagstype adpcm\_d} benchmark.}
%    \label{fig:adpcm_d-cfg-instr}
%\end{figure}

While the \OptProf has a maximum overhead of near 60\%, the \WCRelax-\textit{5}\% has an overhead of
less than 16\% for all benchmarks, with the \WPRelax-\textit{5}\% reaches an average overhead of only 5\%,
compared to an average overhead of 12\% for the optimal profiling.

\begin{figure*}[t!]
    \centering
    \includegraphics[width=\textwidth]{figs/overhead-O3.pdf}
    \caption{Overhead of the instrumentations compiled with {\flagstype -O3}.}
    \vspace{-3mm}
    \label{fig:overhead-O3}
\end{figure*}
\begin{figure*}[t!]
    \centering
    \includegraphics[width=\textwidth]{figs/error-O3.pdf}
    \caption{Dynamic error of the work profiling averaged over the 1000 inputs, after relaxing the number of probes.}
    \vspace{-3mm}
    \label{fig:error-O3}
\end{figure*}

The most critical case, amongst the evaluated benchmarks, happens with the
\texttt{adpcm\_d} program, with an overhead of about 59\% for the optimal
instrumentation.
With a threshold of only 5\%, the \WCRelaxLower relaxation reduces
the overhead by about 75\% over the \OptProf,
and the \WPRelax-5\% reduces it by about 80\%.
This benchmark has a single hot function
which consists mainly of a single hot loop with several branches inside it. %, as depicted by Figure~\ref{fig:adpcm_d-cfg-instr}.
Most of the improvement achieved by both strategies comes from removing only two probes from the hot loop. The two relaxed probes were
placed in branches with a high probability of being taken, but with a small contribution to the total amount of
work measured in DAG for this loop (both probes with an static error of about 1.3\%). Because these probes were placed in basic blocks with
a considerable execution frequency, removing the probes resulted in a significant reduction in runtime overhead.

A few of the benchmarks present abnormal regressions in their execution time.
This increase in the profiling overhead is counter-intuitive because the
relaxation algorithms only reduces the number of instrumented probes,
without changing their placement.
We suggest some reasons for these abnormal regressions.
By analyzing these cases, we noticed that some of the frequently executed
probes could not be removed by the relaxation strategies with small thresholds.
Another important aspect to point out is that we also observed
differences among the final CFGs after applying the \texttt{-O3} optimization.
Although relaxation does not directly alter the CFG, it is worth mentioning that
there are optimizations that make use of LLVM's built-in target-based cost model,
which means that adding or removing instructions from basic blocks may affect
decisions during transformations of the code, which may end up changing the CFG
in different ways depending on the instrumentation applied.
%For example, some of the optimizations that make use of this built-in cost model
%and also alter the CFG are: loop-unrolling, function inlining, and
%simplifications of the CFG.
%These changes in the program's CFGs may affect other compiler optimizations and
%also possibly in hardware performance, for example affecting the instruction
%caching.

Finally, as shown in Figure~\ref{fig:error-O3}, these improvements can be
obtained while incurring very little dynamic error.
This result confirms the overly conservative aspect of the \WCRelaxLower relaxation,
while the \WPRelaxLower relaxation is significantly less conservative.
Although the error bound is not guaranteed by the \WPRelaxLower relaxation,
its dynamic error was always below the 5\% thresholds.
%due to the use of perfect profiling information.

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figs/speedups.pdf}
    \caption{Speedups obtained from the final optimization sequence selected by the online {\itercomp}.
	         The speedups reported for each benchmark represents the average speedup across their complete 1000 input datasets.}
    \label{fig:speedups}
\end{figure*}

In this section we evaluate the effectiveness of our work efficiency metric.
We perform online {\itercomp} using all five profiling strategies, namely
\OracleRM, \OraclePP,\OptProf,\WCRelax-5\%, and \WPRelax-5\%.
Because the \OracleRM measures actual speedup, it provides reference results for
assessing the effectiveness of the work-efficiency metric.
For the relaxation strategies, we evaluate only those with a 5\% threshold, since
they offered better overhead reductions for little dynamic errors in the measurement.

>>>>

@@

\begin{figure*}[t!]
    \centering
    \includegraphics[width=\textwidth]{figs/overhead-O3.pdf}
    \caption{Instrumentation overhead for each benchmark averaged over all inputs, when compiled with {\flagstype -O3}.}
    \vspace{-3mm}
    \label{fig:overhead-O3}
\end{figure*}

\begin{figure*}[t!]
    \centering
    \includegraphics[width=\textwidth]{figs/error-O3.pdf}
    \caption{Dynamic error of the work profiling averaged over the 1000 inputs, after relaxing the number of probes.}
    \vspace{-3mm}
    \label{fig:error-O3}
\end{figure*}

@@

First, we evaluate the runtime overhead introduced by the work efficiency profiling. A high overhead would impact the user's experience
and would make our approach impractical. We take the original and instrumented versions of each benchmark, we compile them using the
baseline \texttt{-O3} setting, and execute them with all 1,000 inputs. The difference in their runtimes, for the same input, is the
instrumentation overhead. Figure~\ref{fig:overhead-O3} shows the average overhead for each benchmark and work profiling strategy.

\OptProf typically slows the application down by 10\%, noticeably affecting the user experience, while in the worst case its overhead
reaches 59\%. It is clearly unsuitable for online work profiling. \WCRelax and \WPRelax fare better, especially when used with a higher
5\% threshold. \WPRelax-\textit{5}\%, in particular, has an average overhead of only 5\%, 13\% in its worst case.

For almost all the profiling strategies, their worst case is \texttt{adpcm\_d}. Still, compared to the 59\% overhead of the \OptProf,
\WCRelax-\textit{5}\% and \WPRelax-\textit{5}\% incur only about a quarter of that overhead. This benchmark has a single hot function
consisting mainly of a single hot loop with several branches inside it. Most of the overhead comes from two probes which are placed in
very frequently executed blocks but contribute little to the total work of the loop. The maximum possible error caused by removing either
of the probes, based on the analysis of the DAG, is about 1.3\%. Both relaxation strategies identify this opportunity and remove the probes.

Some benchmarks display unexpectedly higher overheads under the relaxation strategies. This is counter-intuitive because relaxation only
reduces the number of instrumented probes without changing their placement. By analyzing these cases, we noticed that we could not remove
the most frequently executed probes, so relaxation had little positive impact. At the same time, we removed some probes which were
infrequently executed but were taken into consideration by the LLVM optimization heuristics. In some cases, removed probes caused different
and suboptimal decisions to be made, increasing the runtime.

Figure~\ref{fig:error-O3} shows that profiling with relaxation incurs very little dynamic error. It also confirms that the \WCRelaxLower
relaxation is overly conservative, while the \WPRelaxLower relaxation achieves a better trade-off. Although the error bound is not
guaranteed by the \WPRelaxLower relaxation, its dynamic error is always below the 5\% threshold. For both relaxation strategies the 2\%
threshold is too conservative to be useful. The rest of the paper only evaluates the 5\% threshold.

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figs/speedups.pdf}
    \caption{Speedups obtained from the final optimization sequence selected by the online {\itercomp}.
	         The speedups reported for each benchmark represents the average speedup across their complete 1000 input datasets.}

    \FIXME{Sort  the bars according to speedup, i.e. Oracle should be the last bars.}
    \label{fig:speedups}
\end{figure*}

In this section we evaluate the effectiveness of our work efficiency metric.
We perform online {\itercomp} using all five profiling strategies, namely
\OracleRM, \OraclePP,\OptProf,\WCRelax-5\%, and \WPRelax-5\%.
Because the \OracleRM measures actual speedup, it provides reference results for
assessing the effectiveness of the work-efficiency metric.
